Evolution of ORM Systems: Formal Foundations

### 2013-02-02 doc/description.tex
equals for classes, associations and properties

@@

\paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used. It can be used in case structure of both classes (respectively tables) is equals. The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
\begin{align*}
&	\mathbf{Merge} : Label \times Label \times Label \times State \rightarrow State \\
&	\Phi(s, Merge(l_1, l_2, l_n) = \\
& \;\;\;\;\;\; = \inference{
		\exists c_1, c_2 \in classes(application(s)) \wedge 
		equals(c_1, c_2)}{\begin{gathered}
			state(application(s)) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
			\circ (\forall c \in classes (application(s)), a \in associations(c) \\
			reference(a) = c_1 \vee reference(a) = c_2 : \\
			reference(a) = l_n) \\
			\circ renClass(l_1) \circ renClass(l_2) \end{gathered}}
 \\\\
&	\Psi(s, merge(l_1, l_2, l_n)) = \\
& \;\;\;\;\;\; = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
\forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
\bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(select(name(f), k)) = pkValue(r) : \\value(select(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
 \end{gathered}} \\\\
 & l_1, l_2, l_n \in Label, s \in State
\end{align*}

>>>>

\paragraph{Comparison of Structures} During the evolution there occur situations when the structure of two classes has to be compared.
\begin{align*}
& \mathbf{equals} : Class \times Class \rightarrow Boolean \\
& equals(c_1, c_2) = \begin{cases}
 	\inference{|properties(c_1)| = |properties(c_2)| \\ \wedge
 	\forall p \in properties(c_1) \exists r \in properties(c_2): \\ equals(p, r) \\
 	|associations(c_1)| = |associations(c_2)| \\
 	\wedge \forall a \in associations(c_1) \exists b \in asccosiations(c_2): \\ equals(a,b) \\ }{true}\\\\
 false
 \end{cases}\\\\
& c_1, c_2 \in Class
\end{align*}
\begin{align*}
& \mathbf{equals} : Property \times Property \rightarrow Boolean \\
& equals(p_1, p_2) = \begin{cases}
	\inference{name(p_1) = name(p_2) \\
	\wedge mandatory(p_1) = mandatory(p_2) \\
	\wedge cardinality(p_1) = cardinality(p_2)}{true}
 \\\\ false
 \end{cases}\\\\
& p_1, p_2 \in Property
\end{align*}
\begin{align*}
& \mathbf{equals} : Association \times Association \rightarrow Boolean \\
& equals(a_1, a_2) = \begin{cases}
	\inference{name(a_1) = name(a_2)\\
	\wedge reference(a_1) = reference(a_2) \\
	\wedge startCardinality(a_1) = startCardinality(a_2) \\
	\wedge endCardinaility(a_1) = endCardinality(a_2)
	}{true}
 \\\\ false
 \end{cases}\\\\
& a_1, a_2 \in Association
\end{align*}

@@

\paragraph{Merge of Classes} There occur situations when there is a need to merge two classes and the inlining transformation cannot be used because of name collisions among properties in such a case the $merge$ transformation can be used. It can be used in case structure of both classes (respectively tables) are equal. The transformation creates new rows in the new table - the primary keys from origin instances are replaced by a new ones.
\begin{align*}
&	\mathbf{Merge} : Label \times Label \times Label \times State \rightarrow State \\
&	\Phi(s, Merge(l_1, l_2, l_n) = \\
& \;\;\;\;\;\; = \inference{
		\exists c_1, c_2 \in classes(application(s)) \wedge 
		equals(c_1, c_2)}{\begin{gathered}
			state(application(s)) \\ \circ addClass(cl(ln, properties(c_1), associations(c_1)) \\
			\circ (\forall c \in classes (application(s)), a \in associations(c) \\
			reference(a) = c_1 \vee reference(a) = c_2 : \\
			reference(a) = l_n) \\
			\circ renClass(l_1) \circ renClass(l_2) \end{gathered}}
 \\\\
&	\Psi(s, merge(l_1, l_2, l_n)) = \\
& \;\;\;\;\;\; = \inference{\exists t_1, t_2 \in tables(database(s)) : \\ name(t_1) = l_1 \wedge name(t_2) = l_2 \\ \wedge columns(t_1) = columns(t_2)\\
\forall f \in (foreignKeys(t_1) \cup foreignKey(t_2)) \\ \setminus (foreignKey(t_1) \cap foreignKey(t_2)) \\ : NOTNULL \not \in constraints(f)  }{\begin{gathered}
	state(database(s)) \\ \bullet addTable(l_n, pk("Id", s_1), columns(t_1), foreignKeys(t_1)) \\ 
\bullet (\forall r \in instances(t_1, database(s)) \cup instances(t_2, database(s)) :  \\ n =  row(l_n, next(seq(key(l_n))), pairs(r)), insertRow(l_n, n) \\ 
\bullet (\forall k \in data(database(s)), f \in foreignKeys(k) \\ \wedge reference(f) = reference(r) \\ \wedge value(select(name(f), k)) = pkValue(r) : \\value(select(name(f), k)) = pkValue(n))) \\  \bullet dropTableAndData(l_1) \bullet dropTableAndData(l_2)
 \end{gathered}} \\\\
 & l_1, l_2, l_n \in Label, s \in State
\end{align*}
